Assignment 2

### 2019-06-29 ex2/report/report.tex
Updated report

If we denote by $f^{i}$ the marginal distribution of the feature space which retains all features except the $i$th feature $x_i$, the PDP for $x_i$ feature can be expressed as
\begin{equation}
\text{PDP}_i(x) = \mathbb E_{\boldsymbol X}\left(Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n)\right) % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},x,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
\end{equation}
Empirically, we can approximate the distribution of the feature space by the observed samples. Hence, at a given point $x$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.

\subsection{Individual Conditional Expectation}
The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case

@@

For a feature $X_i$ the ALE plot ALE$_i(x)$ can be defined differentially as

\begin{equation}
\frac{d}{dx} \text{ALE}_i (x) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_x} Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n) | X_i=x\right)
\end{equation}

@@

 Table~\ref{tab:logreg_coeff} displays the coefficients for 
\subsection{Interpretation}
\section{Implementing a Backdoor}

>>>>

If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the PDP for the $i$th feature $X_i$ can be expressed as
\begin{equation}
\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) . % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
\end{equation}
Empirically, we can approximate the distribution of the feature space by the distribution of observed samples. Hence, at a given point $x,$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.

\subsection{Individual Conditional Expectation}
The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well. Hence, by averaging information gets lost and wrong explanations might be the consequence.

For Individual Conditional Expectation (ICE) plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples, i.e.

\begin{equation}
\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
\end{equation}

@@

For the $i$th feature $X_i$, the ALE plot ALE$_i(y)$ can be defined differentially as

\begin{equation}
\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_y} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
\end{equation}

@@

 Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
\subsection{Interpretation}
\section{Implementing a Backdoor}
Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power. In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.

If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.

For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the traning models.

Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors. On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.

In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor. The TTL is used for mitigating problems due to routing loops in IP networks. In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow. In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.

Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.  The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
